---
title: "The negative impact of data analysis on human rights."
Author: HengMa
format: pdf
editor: visual
---

```{r setup, fig.pos = "!H", include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(dplyr)
library(tidyr)
library(opendatatoronto)
library(knitr)
library(patchwork)
library(lubridate)

```


```{r}
packagedata <- search_packages("race")
toronto_gender <-
  list_package_resources("police-race-and-identity-based-data-use-of-force") |>
  filter(name == 
           "Gender Composition") |>
  get_resource()
write_csv(
  x = toronto_gender,
  file = "toronto_gender.csv"
)
knitr::kable(toronto_gender, "latex", booktabs = TRUE) %>% 
  kableExtra::kable_styling(font_size = 7)
head(toronto_gender)
```

##Introduction

As a city with a significant influx of population, Toronto has immigrants making
up more than half of its population, and this is related to the city’s persistently
high crime rate. Many articles illustrate that ethnic diversity is one of the factors contributing to the high crime rate, but considering the complexity of society as a whole, it’s difficult to definitively say that race leads to violence. This paper, by further collecting data from the Open Data Portal of the City of Toronto, aims
to analyze the problem and trends by visualizing data related to the least violent
profession, police, and its association with violence.

##Varialbes 

This data set is divided into four different categories: Type of Incident, Gender of People
Involved, Perceived Race of People Involved, and Incident Count. The Type of Incident is used
to determine whether the violence was recorded by someone else or used in an enforcement
action. Since our focus is on the role of the police and whether the use of force is reactive or
proactive does not affect the data analysis, this will be cleaned out later. Gender represents
the gender of the police officer involved, and like the previous, will not be referenced. Race
and Count, as the most crucial data points for discussion, will be more explicitly laid out.


##Revelence

This quote raises a significant concern about the relationship between data interpretation and human rights, emphasizing the inherent risk of selectively using data to reinforce existing societal biases and prejudices. Historically, various situations involving minimal or misunderstood data have been used to justify the oppression and marginalization of specific groups. This practice not only compromises the objectivity of the data but also poses a severe threat to fundamental human rights principles. For instance, data obtained from the Toronto database, ostensibly for studying the frequency of police violence and its potential direct correlation with race, can be misused. Conversely, this data might be employed to assert that certain races inherently possess more violent tendencies, leading to erroneous perceptions about specific ethnic groups. This misinterpretation could result in severe human rights issues, as it fosters discrimination and perpetuates harmful stereotypes.

##Conclusion

Human rights are based on treating everyone equally and with respect. When data analysis supports racial prejudice or unfair treatment, it goes against these important principles. For example, with the Toronto database, if the data wrongly points to some races being more violent, it can make racial problems worse. This could lead to unfair treatment of people because of their race. Such misuse of data can change how society sees these groups and may lead to biased government policies and police actions, which is not fair and goes against human rights.
